The Spine of American Legal Practice

### 2016-06-25 article/article.Rmd
Minor fixes

---
author:
- Kellen Funk
- Lincoln Mullen
documentclass: article
fontsize: 12pt
output:
  pdf_document:
    fig_caption: True
    number_sections: False
title: The Spine of American Legal Practice
...

@@

After David Dudley Field and two other lawyers were appointed to the procedural commission, their reports made clear that they favored the French model but understood political realities would hold them to the English model. From 1847 to 1850, the commissioners made five reports to the legislature knowing they had no power to keep legislators from amending their code or even defeating it altogether. Each time they reminded legislators that "public opinion had issued its mandate in the most imposing form" of a constitutional decree. The constitution, together with the legislative act appointing the commission, "gave the commissioners instructions so precise, as to leave them no discretion, if they had desired it, \[and\] promised them therefore in advance, so long as they obeyed those instructions, the concurrence and co-operation of all departments of the government."[^16] Although the theory of codification made it a democratic enterprise, in practice Field---an ardent Jacksonian---insisted the democratic legislature ought to defer to the expertise of the commissioners.

@@

Despite Field's arguments that a mere procedure code was democratically unproblematic, his efforts were not entirely successful in New York. The commission submitted a draft of its main reforms in 1848, emphasizing that this first code was "but a report in part." New York's legislators enacted the partial code with little amendment, some legislators repeating Field's view that the constitution obligated them to accept the code.[^25] But when the commissioners submitted their final draft in 1849, the Assembly judiciary committee balked, directly disputing the commissioners' claims that procedure was merely the machinery of the law. They wrote that "provisions for rights and for the mode of pursuing remedies, insensibly run into each other," and observed that distinctions between substance and procedure were difficult to draw in practice. In fact, "the parts of a system of law are more dependent on each other, and run into and affect each other, infinitely more than any machine of human contrivance." They therefore suspected the commissioners' forthcoming code of criminal procedure would include all of the criminal law as well, "as they seem to understand practice and pleadings to include all the law upon a given subject." That being the case, the committee wondered whether they should "place in \[the commissioners\] a blind and implicit confidence that shall commit to their discretion the peace and property, the personal liberty and the lives of those who sent *us* here to make laws for them?"[^26]

@@

Curating a corpus to answer a specific question is an important difference from the way that digital history often describes itself. Speaking in broad terms, digital historical work---like all historical work---can begin either with a set of sources or a question. Digital history is often conceived of as working with very large datasets such as the Google Books or Hathi Trust corpora. These large corpora are sometimes called "big data"---though, it must be emphasized, almost never by digital historians who actually work with them---and are used as the basis of a "distant reading." We think that working with these large corpora is worthwhile because it is salutary for historians to have their research questions shaped by the broadest possible contexts. Nevertheless, it is not apparent that digital historians can readily move from these omnibus corpora to answering the specific research questions that animate historical fields. In this article we demonstrate an alternative approach, which we might unimaginatively label "medium data." Although the amount of legislation that governed American civil practice is impressive---every state amended and re-enacted procedure statutes nearly every decade---a corpus of procedural legislation requires some computational sophistication but is far from needing the computational techniques derived for truly big data. We have gathered a large but narrowly constrained corpus centered on solving a well-defined research question. This corpus is large enough that digital historical methods provide results that a scholar could not obtain through traditional methods, but sufficiently circumscribed so as to directly address a discipline- and field-specific question.[^31]

@@

Even later New York codes can be considered a separate family. In 1876 a New York commission produced a new code attempting to consolidate all the statutory amendments subsequent to the 1851 Field Code. David Dudley Field was upset by the changes introduced in this revised code, called the Throop Code. A count by a "friend" of Field's found that only three sentences of the Field Code had carried over word-for-word into the Throop Code. Nevertheless, many of the changes were purely formal, and Field speculated that these changes had been made only "so that \[the code\] should not appear to be the same thing as before." With respect to Mr. Field or his "friend," we were able to find that the connection was between the Throop Code and the Field Code was somewhat stronger than he thought, and the Throop Code was indeed a descendant, if much revised, from the Field Code.

@@

As in postbellum North Carolina, establishing a flow of credit through the remedial system became a leading priority of western lawyers. While the so-called new western history has shed significant light on neglected topics of Indian dispossession and environmental management, it has often done so by leaving out of view matters of political economy, a staple of the old western history. As one work in that older tradition argued, "Debt collection was the central part of law practice for the \[western\] bar and remained a key part of private practice throughout the century."[^60] On that understanding, one western lawyer succinctly summarized the difference between the code and the common law as "whether a merchant had better try to collect a \$500 note or burn it up." Tiring of all the focus on creditors' remedies, one legislator observed that he "never knew one of these professionals who undertook to write up the beauties of the New York code, ... that he did not also break out somewhere with 'take for instance the case of an action on a promissory note,' as though the collection of notes was about all there could be any law needed for."[^61]

>>>>

---
author:
- Kellen Funk
- Lincoln Mullen
documentclass: article
fontsize: 12pt
output:
  pdf_document:
    fig_caption: true
    number_sections: false
title: The Spine of American Legal Practice
...

@@

After David Dudley Field and two other lawyers were appointed to the procedural commission, their reports made clear that they favored the French model but understood political realities would hold them to the English model. From 1847 to 1850, the commissioners made five reports to the legislature knowing they had no power to keep legislators from amending their code or even defeating it altogether. Each time they reminded legislators that "public opinion had issued its mandate in the most imposing form" of a constitutional decree. The constitution, together with the legislative act appointing the commission, "gave the commissioners instructions so precise, as to leave them no discretion, if they had desired it, [and] promised them therefore in advance, so long as they obeyed those instructions, the concurrence and co-operation of all departments of the government."[^16] Although the theory of codification made it a democratic enterprise, in practice Field---an ardent Jacksonian---insisted the democratic legislature ought to defer to the expertise of the commissioners.

@@

Despite Field's arguments that a mere procedure code was democratically unproblematic, his efforts were not entirely successful in New York. The commission submitted a draft of its main reforms in 1848, emphasizing that this first code was "but a report in part." New York's legislators enacted the partial code with little amendment, some legislators repeating Field's view that the constitution obligated them to accept the code.[^25] But when the commissioners submitted their final draft in 1849, the Assembly judiciary committee balked, directly disputing the commissioners' claims that procedure was merely the machinery of the law. They wrote that "provisions for rights and for the mode of pursuing remedies, insensibly run into each other," and observed that distinctions between substance and procedure were difficult to draw in practice. In fact, "the parts of a system of law are more dependent on each other, and run into and affect each other, infinitely more than any machine of human contrivance." They therefore suspected the commissioners' forthcoming code of criminal procedure would include all of the criminal law as well, "as they seem to understand practice and pleadings to include all the law upon a given subject." That being the case, the committee wondered whether they should "place in [the commissioners] a blind and implicit confidence that shall commit to their discretion the peace and property, the personal liberty and the lives of those who sent *us* here to make laws for them?"[^26]

@@

Curating a corpus to answer a specific question is an important difference from the way that digital history often describes itself. Speaking in broad terms, digital historical work---like all historical work---can begin either with a set of sources or with a question. Digital history is often conceived of as beginning with very large datasets such as the Google Books or Hathi Trust corpora. These large corpora are sometimes called "big data"---though, it must be emphasized, almost never by digital historians who actually work with them---and are used as the basis of a "distant reading." We think that working with these large corpora is worthwhile because it is salutary for historians to have their research questions shaped by the broadest possible contexts. Nevertheless, it is not apparent that digital historians can readily move from these omnibus corpora to answering the specific research questions that animate various historical fields. In this article we demonstrate an alternative approach, which we might unimaginatively label "medium data." Although the amount of legislation that governed American civil practice is impressive---every state amended and re-enacted procedure statutes nearly every decade---a corpus of procedural legislation requires some computational sophistication but is far from needing the computational techniques derived for truly big data. We have gathered a large but narrowly constrained corpus centered on solving a well-defined research question. This corpus is large enough that digital historical methods provide results that a scholar could not obtain through traditional methods, but sufficiently circumscribed so as to directly address a discipline- and field-specific question.[^31]

@@

Even later New York codes can be considered a separate family. In 1876 a New York commission produced a new code attempting to consolidate all the statutory amendments subsequent to the 1851 Field Code. David Dudley Field was upset by the changes introduced in this revised code, called the Throop Code. A count by a "friend" of Field's found that only three sentences of the Field Code had carried over word-for-word into the Throop Code. Nevertheless, many of the changes were purely formal, and Field speculated that these changes had been made only "so that [the code] should not appear to be the same thing as before." With respect to Mr. Field or his "friend," we were able to find that the connection was between the Throop Code and the Field Code was somewhat stronger than he thought, and the Throop Code was indeed a descendant, if much revised, from the Field Code.

@@

As in postbellum North Carolina, establishing a flow of credit through the remedial system became a leading priority of western lawyers. While the so-called new western history has shed significant light on neglected topics of Indian dispossession and environmental management, it has often done so by leaving out of view matters of political economy, a staple of the old western history. As one work in that older tradition argued, "Debt collection was the central part of law practice for the [western] bar and remained a key part of private practice throughout the century."[^60] On that understanding, one western lawyer succinctly summarized the difference between the code and the common law as "whether a merchant had better try to collect a \$500 note or burn it up." Tiring of all the focus on creditors' remedies, one legislator observed that he "never knew one of these professionals who undertook to write up the beauties of the New York code, ... that he did not also break out somewhere with 'take for instance the case of an action on a promissory note,' as though the collection of notes was about all there could be any law needed for."[^61]
